### 2014-03-10 pecomplexity.tex
new abstract

\documentclass[%
 reprint,
%superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%showpacs,preprintnumbers,
nofootinbib,
nobibnotes,
%bibnotes,
 amsmath,amssymb,
% aps,
%pra,
%prb,
%rmp,
%prstab,
%prstper,
%floatfix,
]{revtex4-1}
%\documentclass{article}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
%\geometry{letterpaper}                   % ... or a4paper or a5paper or ...
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parIEEEabrvskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\usepackage{amssymb}
%\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage{amsthm}
\usepackage{xspace}
\usepackage{pifont} % check and x marks
\usepackage{color}

\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

\newcommand{\gcc}{{\tt 403.gcc}\xspace}
\newcommand{\svd}{{\tt dgesdd}\xspace}
\newcommand{\svdone}{{\tt dgesdd}$_1$\xspace}
\newcommand{\svdtwo}{{\tt dgesdd$_2$}\xspace}
\newcommand{\svdthree}{{\tt dgesdd$_3$}\xspace}
\newcommand{\svdfour}{{\tt dgesdd$_4$}\xspace}
\newcommand{\svdfive}{{\tt dgesdd$_5$}\xspace}
\newcommand{\svdsix}{{\tt dgesdd$_6$}\xspace}
\newcommand{\naive}{na\"ive}
\newcommand{\row}{{\tt row\_major}\xspace}
\newcommand{\col}{{\tt col\_major}\xspace}
%\newcommand{\color{red}}{{\color[rgb]{1,0,0}}}

%\parskip=3pt

%\usepackage{authblk}
%\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\newtheorem*{mydef}{Definition}

%Version 1
%\author{
%  \IEEEauthorblockN{Joshua Garland}
%  \IEEEauthorblockA{Dept. of Computer Science\\
%    University of Colorado at Boulder\\
%    Colorado, USA\\
%    Email: joshua.garland@colorado.edu}
%  \and
%  \IEEEauthorblockN{Ryan G.~James}
%  \IEEEauthorblockA{Complexity Sciences Center \& Dept. of Physics\\
%    University of California at Davis\\
%    California, USA\\
%    Email: rgjames@ucdavis.edu}
%    \and
%      \IEEEauthorblockN{Elizabeth Bradley}
%         \IEEEauthorblockA{Santa Fe Institute \\
%    New Mexico, USA\\
%    }
%  \IEEEauthorblockA{Dept. of Computer Science\\
%    University of Colorado at Boulder\\
%    Colorado, USA\\
%    Email: lizb@colorado.edu}
%  }

 %\author{
 %  \IEEEauthorblockN{
 %    Joshua Garland\IEEEauthorrefmark{1},
 %    Ryan G.~James\IEEEauthorrefmark{1} and
 %    Elizabeth Bradley\IEEEauthorrefmark{1}\IEEEauthorrefmark{2}}
 %  \IEEEauthorblockA{
 %    \IEEEauthorrefmark{1}Dept. of Computer Science
 %    University of Colorado, Boulder, Colorado 80309-0430 USA\\
 %    Email: joshua.garland@colorado.edu, ryan.james@colorado.edu}
 %  \IEEEauthorblockA{
 %    \IEEEauthorrefmark{2}Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, New Mexico 87501  USA\\
 %    Email: lizb@colorado.edu}
 %  % \IEEEauthorblockA{
 %  %   \IEEEauthorrefmark{3}Complexity Sciences Center \& Physics Dept., University of California, Davis, %California 95616 USA\\
%   %   Email: rgjames@ucdavis.edu}
% }

\begin{document}

\preprint{APS/123-QED}

%\title{Manuscript Title}% Force line breaks with \\
%\thanks{A footnote to the article title}%

\title{
%Determinism, Complexity, and Predictability in Computer Performance\\
%Quantifying structured complexity and its role in predictability: With applications to predicting computer performance.
%\\ 
{\color{green}Quantifying Predictability through Structural Complexity} 
%\\ 
%Quantifying Predictability using Structural Complexity 
\\Decoding Predictability in Complicated Real-Valued Time Series: The balance between complexity and temporal structure
\\Decoding Complicated Time Series: The balance between complexity and temporal structure and the implications for predictability
}

%\author[1]{Joshua Garland %\thanks{joshua.garland@colorado.edu}}
%\author[1]{Ryan James \thanks{ryan.james@colorado.edu}}
%\author[1,2]{Elizabeth Bradley \thanks{lizb@colorado.edu}}
%\affil[1]{Department of Computer Science\\
%  University of Colorado at Boulder\\
%  Colorado, USA
%}
%\affil[2]{Santa Fe Institute\\
%  New Mexico, USA
%}

\author{Joshua Garland}\email{joshua.garland@colorado.edu}
\author{Ryan James}%
 \email{ryan.james@colorado.edu}
\affiliation{%
 Department of Computer Science, University of Colorado at Boulder. Colorado, USA}%
%\collaboration{MUSO Collaboration}%\noaffiliation
\author{Elizabeth Bradley}\email{lizb@colorado.edu}
\affiliation{
 Department of Computer Science, University of Colorado at Boulder. Colorado, USA
}%
\affiliation{
 Santa Fe Institute, New Mexico, USA\\
 %This line break forced% with \\
}%

\date{\today}% It is always \today, today,
             %  but any date may be explicitly specified

\begin{abstract}
{\color{green}From the site:It should be about 5\% of the length of the article, but less than 500 words.} This paper provides insight
into when, why, and how forecast strategies fail when they are applied
to complicated time series by empirically estimating their inherent
complexity. We focus here on the specific example of computer
performance but believe the results apply to a much broader class of
time series. We conjecture that the complexity of traces from these
systems---which results from the inherent dimension, non-linearity,
and non-stationarity, as well as from measurement issues like noise,
aggregation, and finite data length---is \emph{empirically
  quantifiable}.  To validate this conjecture we model 120 different
performance traces using a variety of forecasting models.  We then
compare the empirically estimated \emph{complexity} with the accuracy
of each forecast method.
\begin{description}
\item[Usage]
Secondary publications and information retrieval purposes.
\item[PACS numbers]
May be entered using the \verb+\pacs{#1}+ \pacs{05.45.Tp} command.
Time series analysis (nonlinear dynamics) 05.45.Tp
Information theory, 89.70.-a
	channel capacity in, 89.70.Kn
	communication complexity in, 89.70.Hj
	computational complexity in, 89.70.Eg
	entropy in, 89.70.Cf
	general biological information, 87.10.Vg
	in neuroscience, 87.19.lo

Nonlinear dynamics, 05.45.-a

	89.75.Kd	Patterns (complex systems)

Computational Physics	
95.75.Wx	Time series analysis, time variability
95.75.Pq	Mathematical procedures and computer techniques
07.05.-t	Computers in experimental physics

\item[Structure]
You may use the \texttt{description} environment to structure your abstract;
use the optional argument of the \verb+\item+ command to give the category of each item. 
\end{description}
\end{abstract}

\pacs{Valid PACS appear here 95.75.Wx	}% PACS, the Physics and Astronomy
                             % Classification Scheme.
%\keywords{Suggested keywords}%Use showkeys class option if keyword
                              %display desired
\maketitle

\input{intro.tex}
\input{related.tex}
\input{methods.tex}

\input{model.tex}

\input{wpe.tex}

\input{results.tex}

\input{concl.tex}

\section*{Acknowledgment}
This work was partially supported by NSF grant \#CMMI-1245947 and ARO
grant \#W911NF-12-1-0288.

\section{Appendix}

\begin{table}[h]
  \begin{center}
  \begin{tabular}{|c|c|c|c|c|}
  \hline
            & MASE LMA    & MASE ARIMA &MASE na\"{i}ve   & $l=6$ \\
 \hline

 \col           & $ 0.050 \pm0.002  $ & $0.599  \pm 0.211 $ & $0.571\pm0.002$&  $0.513 \pm 0.003$ \\

\gcc           & $ 1.530\pm 0.021$ & $1.837 \pm0.016 $ & $0.951 \pm 0.001$ & $0.943 \pm 0.001$ \\

\svdone     & $ 0.827\pm 0.076$ & $ 0.714\pm 0.075 $ & $2.676\pm4.328$&  $0.957 \pm 0.016$ \\

 \svdtwo    & $1.279 \pm0.020 $ & $2.163 \pm0.027 $ &  $3.054\pm0.040$ &   $0.846 \pm0.004$ \\

 \svdthree     & $0.619 \pm0.021 $ & $0.713 \pm 0.010 $ & $31.386\pm 0.282$ &  $0.716 \pm 0.006$ \\

 \svdfour     & $ 0.779\pm0.036 $ & $0.979 \pm0.032 $ & $2.661\pm0.074$ & $0.825 \pm 0.008$ \\

 \svdfive     & $ 0.718\pm 0.048 $ & $2.370  \pm 0.051 $ & $20.870 \pm 0.192$&  $0.678 \pm 0.007$ \\

 \svdsix     & $ 0.739\pm 0.068 $ & $ 1.438\pm 0.0610$ & $2.197\pm0.083$&  $0.748 \pm 0.011$ \\
  \hline
  \end{tabular}
  \end{center}
 \label{default}
 \caption{Detailed MASE scores}
 \label{tab:error}
  \end{table}%

\bibliographystyle{aipauth4-1}
\bibliography{bibliofile}

\end{document}