### 2014-03-27 wpe.tex
make WPE section more readable and consistant

\section{Measuring Structural Complexity }\label{sec:meaComplex}
{\color{blue} EDITABLE}

The goal of the section is to describe how we reasonably estimate the entropy of an arbitrary, real-valued time series. We will first detail why this is a highly non-trivial task, and then we will weave a path through which this can in fact be accomplished. This path will draw upon a myriad of tools and results from a variety of fields including time series analysis, dynamical systems, stochastic processes, and chaos.

For the purposes of this paper one can view the Shannon entropy, and in particular its growth rate  with respect to word length (the \emph{Shannon entropy rate}), as a measure of complexity and unpredictability in a time series. A high-entropy rate time series is almost completely unpredictable---and conversely a low entropy rate signal is often quite predictable. This can be made more rigorous: Pesin's relation \cite{pesin77} states that in chaotic dynamical systems, the Kolmogorov-Sanai (KS) entropy is equal to the sum of the positive Lyapunov exponents, $\lambda_i$. The Lyapunov exponents directly quantify the rate at which nearby states of the system will diverge with time: $\left| \Delta x(t) \right| \approx e^{\lambda t} \left| \Delta x(0) \right|$. The faster the divergence, the larger the entropy and the more difficult prediction becomes. Further, the KS entropy is defined as the supremum of the Shannon entropy rates of all partitions. The partition which achieves this supremum is known as a \emph{generating partition}.

From a different point of view, we can consider the information (as measured by the Shannon entropy) contained in a single observable, say the \emph{present}, of the system. This information can be partitioned in to two components: the information shared with past observations, the mutual information between the past and present, and the information in the present which is not contained in the past, the conditional entropy of the present given the past. The first part is known as the \emph{redundancy}, as it is information in the present which is also in the past, and the second part is the aforementioned \emph{Shannon entropy rate}. In theory, the more redundancy in a signal the more predictable it is. The difference between prediction methods is the types of redundancy (linear, non-linear, etc) they can detect and make use of.

Utilizing Shannon entropy rate as a measure of temporal complexity is not a new idea \cite{Shannon1951, mantegna1994linguistic}, but these approaches required categorical data: $x_t \in \mathcal{S}$ for some finite or countably infinite \emph{alphabet} $\mathcal{S}$, whereas data taken from real-world systems is effectively\footnote{Measurements from real devices are always discrete though typically of very high resolution, for example \texttt{double} precision. Though these values are discete, for the purposes of entropy calculations they are effectively a continuum.} real-valued.  To get around this, one must discretize the data --- typically achieved by binning.  Unfortunately, this is rarely a good solution to the problem, as the binning of the values introduces an additional dynamic on top of the intrinsic dynamics whose entropy is desired\cite{bollt2001}. The field of symbolic dynamics studies how to discretize a time series in such a way that the intrinsic behavior is not perverted, but these methods are fragile in the face of noise and furthermore require understanding of the underlying system, which defeats the purpose of measuring the entropy to quantify predictability in the first place.

Bandt and Pompe introduced the \emph{permutation entropy} (PE) as a ``natural complexity measure for time series'' \cite{bandt2002per}. Permutation entropy employs a method of discretizing real-valued time series that follows the intrinsic behavior of the system under examination.  Rather than looking at the statistics of sequences of values, as is done when computing the Shannon entropy, permutation entropy looks at the statistics of the \emph{orderings} of sequences of values using ordinal analysis. Ordinal analysis of a time series is the process of mapping successive time-ordered elements of a time series to their value-ordered permutation of the same size.  By way of example, if $(x_1, x_2, x_3) = (9, 1, 7)$ then its \emph{ordinal pattern}, $\phi(x_1, x_2, x_3)$, is $231$ since $x_2 \leq x_3 \leq x_1$.  This method has many features; among other things, it is generally robust to observational noise and requires no knowledge of the underlying mechanisms.

\begin{mydef}[Permutation Entropy]

  Given a time series $\{x_t\}_{t = 1,\dots,T}$. Define $\mathcal{S}_\ell$ as all $\ell!$ permutations $\pi$ of order $\ell$. For each $\pi \in \mathcal{S}_\ell$ we determine the relative frequency of that permutation occurring in $\{x_t\}_{t = 1,\dots,T}$:
  \begin{align*}
    P(\pi) = \frac{\left|\{t|t \leq T-\ell,\phi(x_{t+1},\dots,x_{t+\ell}) = \pi\}\right|}{T-\ell+1}
  \end{align*}
  Where $P(\pi)$ quantifies the probability of an ordinal and $|\cdot|$ is set cardinality. The \emph{permutation entropy} of order $\ell \ge 2$ is defined as
  \begin{align*}
    H(\ell) = - \sum_{\pi \in \mathcal{S}_\ell} P(\pi) \log_2 P(\pi)
  \end{align*}

\end{mydef}

Notice that $0\le H(\ell) \le \log_2(\ell!)$ \cite{bandt2002per}.  With this in mind, it is common in the literature to normalize permutation entropy as follows: $\frac{H(\ell)}{\log_2(\ell!)}$.  With this convention, ``low'' PE is close to 0 and ``high'' PE is close to 1. Finally, it should be noted that the permutation entropy has been shown to be identical to the Kolmolgorov-Sinai entropy for many large classes of systems \cite{amigo2012permutation}, so long as observational noise is sufficiently small. As mentioned before, this is equal to the Shannon entropy rate of a generating partition of the system. What this normalization allows us to do, then, is approximate the redundancy of a signal by $1 - \frac{H(\ell)}{\log_2(\ell!)}$.

Here we will be utilizing a variation of the permutation entropy, the \emph{weighted permutation entropy} (WPE)~\cite{fadlallah2013}. The weighted permutation entropy attempts to correct for observational noise which is larger than some trends in the data, but smaller than the larger scale features --- for example, a signal that switches between two fixed points with noise about those fixed points. The weighted permutation entropy would be dominated by the switching rather than by the stochastic fluctuation. To accomplish this, the \emph{weight} of a permutation is taken into account:
\begin{align*}
  w(x_{t+1:t+\ell}) = \frac{1}{\ell} \sum_{x_i \in \{x_{t+1:t+\ell}\}}
                      \left( x_i - \bar{x}_{t+1:t+\ell} \right)^2
\end{align*}
where $x_{t+1:t+\ell}$ is a sequence of values $x_{t+1}, \ldots, x_{t+\ell}$, and $\bar{x}_{t+1:t+\ell}$ is the arithmetic mean of those values.

The weighted probability of a permutation is then:
\begin{align*}
  P_w(\pi) = \frac{\displaystyle \sum_{t \le T - \ell} w(x_{t+1:t+\ell}) \cdot \delta(\phi(x_{t:t+\ell}), \pi) }{\displaystyle \sum_{t \le T - \ell} w(x_{t+1:t+\ell})}
\end{align*}
where $\delta(x, y)$ is 1 if $x = y$ and 0 otherwise. Effectively, this weighted probability enhances permutations involved in ``large'' features and demotes permutations which are small in amplitude relative to the features of the time series. The weighted permutation entropy is then:
\begin{align*}
  H_w(\ell) = - \sum_{\pi \in \mathcal{S}_\ell} P_w(\pi) \log_2 P_w(\pi),
\end{align*}
which can also be normalized by dividing by $\log_2(\ell!)$, and will be in all the results of this paper.

In practice, calculating permutation entropy and weighted permutation entropy involves choosing a good value for the word length $\ell$. The primary consideration is that the value be large enough that forbidden ordinals are discovered, yet small enough that reasonable statistics over the ordinals are gathered: e.g.:
\begin{align*}
  \ell = \operatornamewithlimits{argmax}_n \{ T \gtrapprox 100 n! \},
\end{align*}
assuming an average of 100 counts per ordinal is sufficient. In the literature, $3 \le \ell \le 6$ is a standard choice --- generally without any formal justification. In theory, the permutation entropy should reach an asymptote with increasing $\ell$, but that requires an arbitrarily long time series. In practice, what one should do is calculate the \emph{persistent} permutation entropy by increasing $\ell$ until the result converges, but data length issues can intrude before that convergence is reached. Here, we found that $\ell = 6$ was a good choice, because with $\ell \geq 7$ many signals had entropies plummet due to finite data effects and so with a word length of 6 we are capturing as much structure as possible without sacraficing the accuracy of ordinal statistics.

%\subsection{Regime Selection:Detecting complexity shift in a time series}\label{sec:wpeRegime}
%
%{\color{red} this is from a previous paper I wrote, if this is used we need to update the plots to use wpe and windowed wpe instead of standard pe, also not sure how much of this is necessary but added as starting point}
%
%Permutation entropy is a measure of the complexity of time series as a whole, but it can also be used to calculate a dynamical shift in a time series as illustrated in \cite{cao2004det}. To detect dynamical changes in a time series we follow the methods of \cite{cao2004det}. Partition the time series  $\{x_t\}_{t = 1,\dots,T}$ into overlapping\footnote{Following \cite{cao2004det} we use maximal overlapping block. (A window shift of 1.)} (or nonoverlapping) blocks. Calculate $H(n)$ for each block, and treat $H(n)$ as a function of time.  According to \cite{cao2004det} drastic changes in permutation entropy can indicate change in the underlying model.
%
%In addition to introducing this block method, \cite{cao2004det} introduce the use of a \emph{lag} when calculating permutation entropy. This is identical to the lag used in delay coordinate embedding and often the same language is used. Cao et al. describe the process of symbolization in \cite{bandt2002per} as embedding the time series in an $m$ dimensional symbol space. When we symbolize a scalar time series,$\{x_t\}_{t = 1,\dots,T}$, we first embed the scalar time series into $m$ dimensional vectors before matching this to a symbol in $\mathcal{S}_m$. So pre-symbolization our time series changes from scalar to objects of the form:  $$[x(i),\,x(i+1),\,\dots,x(i+(m-1)]$$ This vector is then mapped by $\phi$ to a symbol in $\mathcal{S}_m$. Cao et al. argue that we can view this as embedding the time series in an $m$ dimensional space with lag 1. Cao et al. then state ``Since in practice the optimal $L$[lag] may be different from 1, we shall present the idea for any $m$[embedding dimension] and $L$."\footnote{No further explanation is given about why this is the case.} To not confuse notation we define $\tau$ to be the lag parameter. The process of symbolization then starts with transforming the scalar time series into a set of vectors of the form $$[x(i),\,x(i+\tau),\,\dots,x(i+(m-1)\tau]$$ where $\tau\ge1$.
%
%\subsubsection{Example: Concatenated Time Series}\label{sec:concatenate}
%As an example of this detection, consider Figure \ref{fig:ts-banded} (a). This is a plot of a concatenated time series comprised of several orbits of the Logistic map with varying parameters as well as intermittent noise. Each concatenated band (either an orbit of the Logistic map or noise) is 50,000 points. The bands of noise are generated with Matlab's \verb|randsample|, using the chaotic bands as the sampling population. Each Logistic map trajectory begins at $x=0.6$. The period-two trajectories have an $R$ of 3.2, the period-three trajectory has an R of 3.838, and the chaotic trajectories have an $R$ of 3.65.
%
%The concatenated time series is comprised of two period-two trajectories($t=0-50,000$ and $t=250,000-300,000$), one period-three trajectory($t=100,000-150,000$), two chaotic trajectories ($t=50,000-100,000$ and $t=200,000-250,000$, and two different bands of noise($t=150,000-200,000$ and $t=300,000-350,000$). Visually the bands of noise and the chaotic trajectory are indistinguishable. If this were an observed time series from a real process, it would be impossible to visually identify which bands were generated deterministically and which were simply noise. If we calculate block permutation entropy as discussed in \cite{cao2004det} and plot this as a function of time it becomes very clear which bands have structure and which do not.
%
%In Figure \ref{fig:ts-banded} (b) we plot the concatenated time series, as well as permutation entropy as a function of time(red, blue and green curves). The red, blue and green curves are the permutation entropy calculated using $m=5$ and $\tau = 1,2,3$ (respectively). Using these curves it is clear that bands 2 and 5 ($t=50,000-100,000$ and $t=200,000-250,000$) are chaotic because they have structure\footnote{Permutation entropy less than one.}. On the other hand, bands 4 and 7 ($t=150,000-200,000$ and $t=300,000-350,000$) have no structure\footnote{Permutation entropy of one.} and are the noise bands. This is in fact the case when comparing to the way the time series was constructed.
%
%The main advantage we can see to the use of lags other than 1 is for analysis of periodicity. This is also the major advantage listed in \cite{cao2004det}. Notice in the period 2 bands the blue curve (permutation entropy calculated with a lag of 2) is zero. This means that with a lag of 2 the period 2 orbit is perfectly structured\footnote{Equivalently, not enough permutations are being realized to impact the calculation of $H(n)$}. If we increase the lag to 3 (the green curve) the two period-two bands have a little bit of structure(as some permutations will be realized) but the period 3 band has zero permutation entropy.
%
%%TSLogistic = [period2Logistic;chaosLogistic;period3Logistic;randomSampleFromMap2;chaosLogistic;period2Logistic;randomSampleFromMap3];
%
%%\begin{figure}
%%  \centering
%%  \subfigure{\includegraphics[width=\columnwidth]{unused-figs/ts-banded-alone}}
%%    {(a)}
%%
%%  \subfigure{\includegraphics[width=\columnwidth]{unused-figs/ts-banded-pe}}
%%  {(b)}
%%  \caption{ (a) The concatenated time series discussed in Section~\ref{sec:concatenate}. (b) Plot of the concatenated time series from Section~\ref{sec:concatenate}, as well as permutation entropy as a function of time(blue,green and red curves). Each permutation entropy curve shown is calculated using $m=5$, however, the red,blue and green curve use a $\tau = 1,2,3$ respectively.}\label{fig:ts-banded}
%%\end{figure}
%
%
%
%
%
%
%and what is it good for with forward pointer to the other sections
%
%
%
%\subsubsection{Example: The transient Logistic map}
%In Section~\ref{sec:concatenate} we discussed a time series which stayed in a parameter regime for long amounts of time and then abruptly changed. As we showed, permutation was a great proxy to detect this change. Another type of parameter change that may occur in a physical system is constant parameter drift, i.e., some parameter slowly changes over time. To explore this type of parameter drift we construct the same time series as was used in \cite{cao2004det}. Iterate the logistic map, starting with $R=2.8$, then at every iteration increase $R$ by $10^{-5}$ until $R=4$. The time series, which can be seen in Figure~\ref{fig:translogistic} appears very similar to the standard bifurcation diagram for the Logistic map. The curves in red and blue are the permutation entropy calculated with $m=5$ and $\tau = 1,2$ (respectively).
%
% As we can see, as bifurcations occur in the dynamics the permutation entropy changes as well. An interesting feature of block calculations are the spikes that occur at bifurcation points. As the block begins to encompass multiple regimes, e.g., fixed point and  period two, permutations are being realized for both regimes. This overlap causes an increase in the time series complexity, and thus a spike in the permutation entropy. The dips you see in the second half correspond to bands of periodicity which occur between the chaos. It is worth noting, that the blue curve (permutation entropy with lag 2) does not go to zero in the period 2 parameter range. This is because this is not a true period 2 orbit but transient that is moving across \emph{several} period 2 orbits, thus it maintains a higher level of complexity, albeit still very low.
%
% This illustrates that permutation entropy can not only distinguish between instant (and drastic) parameter shifts but also very slow and gradual parameter drift.
%
%\begin{figure}[ht]
%  \centering
%  \includegraphics[width=\columnwidth]{unused-figs/tstranslog-perm}
%  \caption{Transient Logistic map time series and permutation entropy curves. Both permutation entropy curves are calculated with $m=5$, the red and blue correspond to $\tau = 1,2$ respectively.}
%  \label{fig:translogistic}
%\end{figure}
%%\begin{enumerate}
%%\item do with henon bifurcation. what happens near the broken chaotic window
%
%%\end{enumerate}
%
%
%The weighted permutation entropy for the \svd program is given in
%Fig.~\ref{fig:wwpe}. To generate this image a window of 5,000 values slid over
%the time series. Within each of those windows, the statistics over words of
%length 4 are computed and the WPE is calculated. The gray bands denote regions
%where the 5,000 value window overlapped visually-distinct regimes. It can be
%seen that the behaviors of the weighted permutation entropy vary between
%regimes. [[I think here it would be good to add a paragraph explaining the windowed WPE was used for regime choices on SVD...emphasizing  that over a time series permutation entropy fluctuates illustrating within a single time series different levels of complexity and predictability exist. Maybe point at some of the predicting predictability papers.]]
%
%%\begin{figure}[htbp]
%%  \centering
%%  \includegraphics[width=1.0\textwidth]{figs/SVD_wwpe}
%%  \caption{[[Joshua: I think adding the colored SVD trace to this would be %good or putting it above this figure]]The weighted permutation entropy of %one run of SVD. The gray bands
%%    are regions where the window overlaps regimes. The window size used is
%%    $5,000 \times 100,000$ instructions and the word length is $4$.}
%%  \label{fig:wwpe}
%%\end{figure}
%
%
%%%FIGURE INTENT
%%This figure shows that entropy changs over time with the signal as well as why the regimes for \svd were chosen. 
%%%%%%%%%%%%%%%%%%
%\begin{figure}[htbp]
%  \centering
%  %\begin{subfigure}{0.3\textwidth}
%  %  \includegraphics[width=0.3\textwidth]{figs/svdipcregimescolored.png}
%  %  \caption{The instructions per cycle of \svd. Each color corresonds to the different regimes as selected by rapid shifts in WPE, as seen in Figure \ref{fig:svd_wwpe}. From left to right each change in color represents a change in regime for 6 regimes in total. }
% %   \label{fig:svd_ts}
% % \end{subfigure}%
%  %\\
% % \begin{subfigure}{\textwidth}
%
% {\color{red}TODO: Put this image back in.}
%%Liz commented this image out temporarily so that her mac doesn't hang
%%when she scrolls past p12
%% We decided to skip the figure completely, since we took out
%% all stuff about PE as regime-shift detection.  
%% [[should add that as future work, though.]]
%  \includegraphics[width=\columnwidth]{figs/SVD_wwpe}
%  \caption{The weighted permutation entropy of one run of SVD. The gray bands are regions where the window overlaps regimes. The window size used is 5,000 $\times$ 100,000 instructions and the word length is $6$. For reference the instructions per cycle of \svd are plotted as a ghost behind this plot. Each color on the ghosted time series corresonds to the different regimes as determined by visual inspection. From left to right each change in color represents a change in regime for 6 regimes in total.}\label{fig:wwpe}
%\end{figure}
%