### 2014-04-02 intro.tex
edits and consistency changes

\section{Introduction  }\label{sec:intro}
%****NOTES START HERE****

%BRAINSTORMING
%\begin{itemize}
%\item Complexity need not be hard to predict (can %point at the simple predictions paper) [[move to %introduction]]
%\item random walk for example is best predicted by %guess what just happend[[move to introduction]]
%\item The kind of complexity present matters, i.e., %that is whether the complexity is structured or not.%[[use here and mention in intro]]

%\item \col brings about the point nicely that some %prediction strategies cannot utilize the processes internal information transfer method. That is a nonlinear internal information transfer system cannot be predicted effectively with a linear strategy. This gives a practitioner leverage on when to give up and when to keep working. [[use in this section as bridge to next section]]

%\end{itemize}

%\begin{it}
%Paragraph on computer performance, including citations to Todd paper
%and summary of the results that indicate that they're deterministic
%nonlinear dynamical systems.  Given that, we should be able to
%predict.  What benefits would accrue if we could do so: power mgmt,
%end world hunger [[this is my primary goal everyday :)]], etc.
%\end{it}
%Things to add to introduction
%\begin{enumerate}

%\item \cmark Make an argument that Computer Performance %is a great testing ground as it gives signals that completely cover the spectrum of complexity \col \dots \gcc

%\item When deterministic structure even complex structure exists that structure can be utilized for prediction.
%\item For noisy real-valued time series distinguishing randomness (WN,RW) complexity from structured nonlinear / chaotic /high period / high dimensional etc complexity is (until now) very hard.
%\subitem for this provide predictions of \gcc and \col side by side and discuss "How can we tell if we did a bad job because the method is inadequate vs the signal being too complex. Lead this into is it possible to tell if there exists structure in a time series to know if we should find a better model or not. Maybe even having 4 predictions. top being ARIMA of the above signals and bottom being LMA of the above signals. Show that one improved and one did not. Is it that we used the wrong method to predict or is it that we simply can't predict the signal better than a random walk due to high levels of internal signal complexity.

%\item\cmark Introduce the two main contributions of the paper which are outlined at the begining of the results section

%\end{enumerate}
%\noindent ***NOTES END HERE****
%
%\noindent ****SECTION START****

Complicated time-series data are ubiquitous in modern scientific
research.  The complexity of these data spans a wide range.  On the
low end of this spectrum are time series that exhibit perfect
predictive structure, i.e, signals whose future values can be
successfully predicted from past values.  This can be viewed as the
output of an underlying process that generates information and/or
transmits it from the past to the future in a perfectly predictable
fashion.  Constant or periodic signals, for example, fall in this
class.  On the opposite end of this spectrum are signals that are what
we call \emph{fully complex}, where the underlying generating process
transmits no information at all from the past to the future.  White
noise and random-walk processes fall in this class.  In fully complex
signals, knowledge of the past gives no insight into the future,
\emph{regardless} of what model one chooses to use.  Signals in the
midrange of this spectrum---e.g., deterministic chaos---pose
interesting challenges from a modeling perspective.  In these signals,
enough information is being transmitted from the past to the future
that an \emph{ideal} model---one that captures the generating
process---can forecast the future behavior of the observed system with
high accuracy.

As a corollary of the undecidability of the halting
problem~\cite{halting-problem}, however, no single forecasting schema
is ideal for all noise-free deterministic
signals~\cite{weigend-book}---let alone all real-world time-series
data sets.  This leads naturally to an important and challenging
question: given a complicated real-valued time series, does there
exist a forecast model that can utilize the information (if any) that
is being transmitted from past to future by the underlying generating
process?  A first step in answering this question is to reliably
quantify where on the complexity spectrum a given time series falls; a
second step is to determine how complexity and predictability are
related.  These are the goals of this paper.

In particular, we wish to determine---without knowing anything about
the generating process---whether a time series is too complex to successfully
predict.  An important practical corollary to this is a strategy for
assessing appropriateness of forecast methods.  If the forecast
produced by a particular method is poor but the time series contains a
significant amount of predictive structure, one can reasonably conclude that that method is inadequate
to the task and that one should seek another method.

The information in an observation can be partitioned into two pieces:
redundancy and entropy generation~\cite{crutchfield2003}.  
\label{page:redundancy}
Our approach exploits this decomposition in order to assess how much
predictive structure is present in a signal---i.e., where it falls on
the complexity spectrum mentioned above.  We define \emph{complexity}
as a particular approximation of Kolmogorov-Sinai
entropy~\cite{KS-entropy}.  That is, we view a random-walk time series
(which exhibits high entropy) as purely complex, whereas a periodic
signal, which exhibits low entropy, is on the low end of the
complexity spectrum.  We argue that an extension of \emph{permutation
  entropy}~\cite{bandt2002per}---a method for approximating the
entropy through ordinal analysis---is an effective way to assess the
complexity of a given time series.  Permutation entropy is ideal for
our purposes because it works with real-valued data and is known to
converge to the true entropy value. Other existing techniques either
require specific knowledge of the generating process or produce biased
values of the entropy~\cite{bollt2001}.

We focus on real-valued, scalar, time-series data that appear
complicated: i.e., not periodic, linear, constant, etc.  We make no
assumptions about the properties of the generating process: whether it
is linear, nonlinear, deterministic, stochastic, stationary,
non-stationary, etc.
To explore the relationship between complexity, predictive structure,
and predictability, we generate forecasts of a variety of time-series
datasets using four different prediction methods, then compare the
accuracy of those predictions to the permutation entropy of the
associated signals.  This results in two primary findings:
\begin{enumerate}
\item The complexity of a noisy real-valued time series is {\sl (i)}
  quantifiable by permutation entropy and {\sl (ii)} correlated with prediction
  accuracy of an appropriate predictor.
\item The way information is generated and processed internally by a
  system plays a crucial role in the success of different forecasting
  schema---and in the choice of which one is appropriate for a given 
time series.
\end{enumerate}
The forecast methods used in this study were chosen to be a
representative sampling of standard prediction strategies, but they do
not, of course, cover that space exhaustively.  Our goal here is an
empirical assessment of the relationship between predictability and
complexity, not formal results about a ``best'' predictor for a given
time series.  From a practitioner's standpoint, it would be useful to
know {\sl a priori} if a time series contains enough predictive structure to
make it worth spending the time and effort searching for a good
forecast method.  It would also be useful to know if a given method is
inadequate---that is, if one could do better and should keep
searching.

%[[Introduce and make it clear why we choose comp. perf. as test bed.]]

For the purposes of this study, we require a broad array of
time-series datasets from across the complexity spectrum.  Rather than
use synthetic data, we study sensor data from computer performance
experiments.  While this is not a common laboratory experiment, it is
a highly appropriate choice here. {\color{red} [[Liz, these are the few sentences I was worried about being too similar but we may have only used them in a grant proposal.]]} Computers are among the most
complex engineered artifacts in current use.  Modern microprocessor
chips contain multiple processing units and multi-layer memories, for
instance, and they use complicated hardware/software strategies to
move data and threads of computation across those resources.  As a
result, the processor and memory loads during the execution of even a
very simple program can evolve in a very complicated fashion.
Figure~\ref{fig:col-ipc} shows a performance trace of a three-line
loop that repeatedly initializes the upper triangle of a matrix in
column-major order.
%
 \begin{figure}[htbp]
    \centering
    \includegraphics[width=\columnwidth]{figs/colshortts}
    % where an .eps filename suffix will be assumed under latex,
    % and a .pdf suffix will be assumed for pdflatex
    \caption{A computer performance trace: the processor load of \col,
      a three-line C program that repeatedly initializes a matrix in
      column-major order, running on an Intel
      i7\textsuperscript{\textregistered}-based machine.}
   \label{fig:col-ipc}
  \end{figure}
%
Although this time series appears to be periodic, it is actually
chaotic~\cite{mytkowicz09}, which places fundamental limits on
predictability.  Changes in the code can radically change this
behavior; indeed, many performance traces exhibit interesting regime
changes as a program moves through the different phases of its
operation.  The resulting traces span the whole range of the
complexity spectrum, from completely predictable to completely
unstructured, making this an ideal testbed for this
study\footnote{Predicting the \emph{state} of a computer, of course,
  would amount to solving the halting problem.  What we are doing here
  is predicting computer \emph{performance}, which does not violate
  the Rice-Shapiro theorem~\cite{hopcroft2007}.}.
% footnote to appease the theoretician in Josh

%The computer systems community has applied a variety of prediction
%strategies to traces like this, most of which employ regression.  An
%appealing alternative builds on the recently established fact that
%computers can be effectively modeled as deterministic nonlinear
%dynamical systems \cite{mytkowicz09}.  This result %implies the
%existence of a deterministic forecast rule for those %dynamics.  In
%particular, one can use \emph{delay-coordinate %embedding} to
%reconstruct the underlying dynamics of computer %performance, then use
%the resulting model to forecast the future values of computer
%performance metrics such as memory or processor loads
%\cite{josh-ida2011}.  In the case of simple microkernels like the one
%that produced the trace in Figure~\ref{fig:ipc}, this deterministic
%modeling and forecast strategy works very well.  In more-complicated
%programs, however, such as numerical software or compilers,
%this forecast strategy---as well as the traditional methods---break
%down some of the time, but work fine others.

%We argue that \emph{permutation entropy}
%\cite{bandt2002per}, a method for measuring the %entropy of a
%real-valued-finite-length time series through ordinal %analysis, is an
%effective way to explore that conjecture.  %{\color{red} Probably remove this and move a%s it is in the experimental methods or point to that section for more details}We study three
%examples---a simple microkernel and two complex %programs: one from the
%SPEC 2006CPU benchmark suite, and one from LAPACK---%running on an Intel i7-based machine.  For
%each program, we calculate the permutation entropy of %the processor
%load (instructions per cycle), then compare that to %the prediction accuracy attainable for
%that trace using a series of deterministic models.

The rest of the paper is organized as follows.
Section~\ref{sec:related} discusses previous results on generating
partitions, local modeling, and error distribution analysis, and
situates this work in that context. Section~\ref{sec:methods} covers
the experimental setup and methods used to collect each time
series. Section~\ref{sec:model} describes the prediction models used
in this study.  In Section~\ref{sec:meaComplex}, we review permutation
entropy, the technique that we use to measure complexity.  In
Section~\ref{sec:results}, we estimate the complexity of each time
series and compare that complexity to the accuracy of predictions
produced by the methods of Section~\ref{sec:model}, operating on that
time series.  In Section~\ref{sec:conc}, we discuss these results and
their implications, and consider future areas of research.