### 2014-04-22 intro.tex
Liz is done rewriting

\section{Introduction  }\label{sec:intro}
%****NOTES START HERE****

%BRAINSTORMING
%\begin{itemize}
%\item Complexity need not be hard to predict (can %point at the simple predictions paper) [[move to %introduction]]
%\item random walk for example is best predicted by %guess what just happend[[move to introduction]]
%\item The kind of complexity present matters, i.e., %that is whether the complexity is structured or not.%[[use here and mention in intro]]

%\item \col brings about the point nicely that some %prediction strategies cannot utilize the processes internal information transfer method. That is a nonlinear internal information transfer system cannot be predicted effectively with a linear strategy. This gives a practitioner leverage on when to give up and when to keep working. [[use in this section as bridge to next section]]

%\end{itemize}

%\begin{it}
%Paragraph on computer performance, including citations to Todd paper
%and summary of the results that indicate that they're deterministic
%nonlinear dynamical systems.  Given that, we should be able to
%predict.  What benefits would accrue if we could do so: power mgmt,
%end world hunger [[this is my primary goal everyday :)]], etc.
%\end{it}
%Things to add to introduction
%\begin{enumerate}

%\item \cmark Make an argument that Computer Performance %is a great testing ground as it gives signals that completely cover the spectrum of complexity \col \dots \gcc

%\item When deterministic structure even complex structure exists that structure can be utilized for prediction.
%\item For noisy real-valued time series distinguishing randomness (WN,RW) complexity from structured nonlinear / chaotic /high period / high dimensional etc complexity is (until now) very hard.
%\subitem for this provide predictions of \gcc and \col side by side and discuss "How can we tell if we did a bad job because the method is inadequate vs the signal being too complex. Lead this into is it possible to tell if there exists structure in a time series to know if we should find a better model or not. Maybe even having 4 predictions. top being ARIMA of the above signals and bottom being LMA of the above signals. Show that one improved and one did not. Is it that we used the wrong method to predict or is it that we simply can't predict the signal better than a random walk due to high levels of internal signal complexity.

%\item\cmark Introduce the two main contributions of the paper which are outlined at the begining of the results section

%\end{enumerate}
%\noindent ***NOTES END HERE****
%
%\noindent ****SECTION START****

Complicated time-series data are ubiquitous in modern scientific
research.  The complexity of these data spans a wide range.  On the
low end of this spectrum are time series that exhibit perfect
predictive structure, i.e, signals whose future values can be
successfully predicted from past values.  Signals like this can be
viewed as the product of an underlying process that generates
information and/or transmits it from the past to the future in a
perfectly predictable fashion.  Constant or periodic signals, for
example, fall in this class.  On the opposite end of this spectrum are
signals that are what we call \emph{fully complex}, where the
underlying generating process transmits no information at all from the
past to the future.  White noise and random-walk processes fall in
this class.  In fully complex signals, knowledge of the past gives no
insight into the future, regardless of what model one chooses to use.
Signals in the midrange of this spectrum---e.g., deterministic
chaos---pose interesting challenges from a modeling perspective.  In
these signals, enough information is being transmitted from the past
to the future that an \emph{ideal} model---one that captures the
generating process---can forecast the future behavior of the observed
system with high accuracy.

As a corollary of the undecidability of the halting
problem~\cite{halting-problem}, however, no single forecasting schema
is ideal for all noise-free deterministic
signals~\cite{weigend-book}---let alone all real-world time-series
data sets.  This leads naturally to an important and challenging
question: given a complicated real-valued time series, does there
exist a forecast model that can leverage the information (if any) that
is being transmitted from past to future by the underlying generating
process?  A first step in answering this question is to reliably
quantify where on the complexity spectrum a given time series falls; a
second step is to determine how complexity and predictability are
related.  These are the goals of this paper.

In particular, we wish to determine---without knowing anything about
the generating process---whether a time series is too complex to successfully
predict.  An important practical corollary to this is a strategy for
assessing appropriateness of forecast methods.  If the forecast
produced by a particular method is poor but the time series contains a
significant amount of predictive structure, one can reasonably conclude that that method is inadequate
to the task and that one should seek another method.

The information in an observation can be partitioned into two pieces:
redundancy and entropy generation~\cite{crutchfield2003}.  
\label{page:redundancy}
Our approach exploits this decomposition in order to assess how much
predictive structure is present in a signal---i.e., where it falls on
the complexity spectrum mentioned above.  We define \emph{complexity}
as a particular approximation of Kolmogorov-Sinai
entropy~\cite{lind95}.  That is, we view a random-walk time series
(which exhibits high entropy) as purely complex, whereas a low-entropy
periodic signal is on the low end of the complexity spectrum.  We
argue that an extension of \emph{permutation
  entropy}~\cite{bandt2002per}---a method for approximating the
entropy through ordinal analysis---is an effective way to assess the
complexity of a given time series.  Permutation entropy is ideal for
our purposes because it works with real-valued data and is known to
converge to the true entropy value. Other existing techniques either
require specific knowledge of the generating process or produce biased
values of the entropy~\cite{bollt2001}.

We focus on real-valued, scalar, time-series data.
% 
% that appears complicated: i.e., not periodic, linear, constant, etc.
% 
We make no assumptions about the properties of the generating process:
whether it is linear, nonlinear, deterministic, stochastic,
stationary, non-stationary, etc.  To explore the relationship between
complexity, predictive structure, and predictability, we generate
forecasts for a variety of time-series datasets using four different
prediction methods, then compare the accuracy of those predictions to
the permutation entropy of the associated signals.  This results in
two primary findings:
\begin{enumerate}
\item The complexity of a noisy real-valued time series is \emph{(i)}
  quantifiable by permutation entropy and \emph{(ii)} correlated with
  prediction accuracy of an appropriate predictor.
\item The way information is generated and processed internally by a
  system plays a crucial role in the success of different forecasting
  schema---and in the choice of which one is appropriate for a given 
time series.
\end{enumerate}
The forecast methods used in this study were chosen to be a
representative sampling of standard prediction strategies, but they do
not, of course, cover that space exhaustively.  Our goal here is an
empirical assessment of the relationship between predictability and
complexity, not formal results about a ``best'' predictor for a given
time series.  From a practitioner's standpoint, it would be useful to
know \emph{ a priori} if a time series contains enough predictive
structure to make it worth spending the time and effort searching for
a good forecast method.  It would also be useful to know if a given
method is inadequate---that is, if an alternative method could do
better and so one should continue searching.

%[[Introduce and make it clear why we choose comp. perf. as test bed.]]

For the purposes of this study, we require a broad array of
time-series datasets from across the complexity spectrum.  Since
forecasting is a real-world problem, we chose to study sensor data
from a computer-performance experiment.  While this is not a common
laboratory experiment, it is a highly appropriate choice here.
Computers are extremely complicated systems.  Modern microprocessor
chips contain multiple processing units and multi-layer memories, for
instance, and they use complicated hardware/software strategies to
maximize performance by moving data and threads of computation across
those resources.  As a result, the processor and memory loads during
the execution of even a very simple program can evolve in a very
complicated fashion.  Figure~\ref{fig:col-ipc} shows a performance
trace of a short program that repeatedly initializes the upper
triangle of a matrix in column-major order.
%
 \begin{figure}[htbp]
    \centering
    \includegraphics[width=\columnwidth]{figs/colshortts}
    % where an .eps filename suffix will be assumed under latex,
    % and a .pdf suffix will be assumed for pdflatex
    \caption{A computer performance trace: the processor load of \col,
      a simple C program that repeatedly initializes a matrix in
      column-major order, running on an Intel
      i7\textsuperscript{\textregistered}-based machine.}
   \label{fig:col-ipc}
  \end{figure}
%
Although this time series appears to be periodic, it is actually
chaotic~\cite{mytkowicz09}, which places fundamental limits on
predictability.  Changes in the code can radically change this
behavior; indeed, many performance traces exhibit interesting regime
changes as a program moves through the different phases of its
operation.  The resulting traces span the whole range of the
complexity spectrum, from completely predictable to completely
unstructured, making them an ideal testbed for this
study\footnote{Predicting the \emph{state} of a computer, of course,
  would amount to solving the halting problem.  What we are doing here
  is predicting computer \emph{performance}, which does not violate
  the Rice-Shapiro theorem~\cite{hopcroft2007}.}.
% footnote to appease the theoretician in Josh

%The computer systems community has applied a variety of prediction
%strategies to traces like this, most of which employ regression.  An
%appealing alternative builds on the recently established fact that
%computers can be effectively modeled as deterministic nonlinear
%dynamical systems \cite{mytkowicz09}.  This result %implies the
%existence of a deterministic forecast rule for those %dynamics.  In
%particular, one can use \emph{delay-coordinate %embedding} to
%reconstruct the underlying dynamics of computer %performance, then use
%the resulting model to forecast the future values of computer
%performance metrics such as memory or processor loads
%\cite{josh-ida2011}.  In the case of simple microkernels like the one
%that produced the trace in Figure~\ref{fig:ipc}, this deterministic
%modeling and forecast strategy works very well.  In more-complicated
%programs, however, such as numerical software or compilers,
%this forecast strategy---as well as the traditional methods---break
%down some of the time, but work fine others.

%We argue that \emph{permutation entropy}
%\cite{bandt2002per}, a method for measuring the %entropy of a
%real-valued-finite-length time series through ordinal %analysis, is an
%effective way to explore that conjecture.  %{\color{red} Probably remove this and move a%s it is in the experimental methods or point to that section for more details}We study three
%examples---a simple microkernel and two complex %programs: one from the
%SPEC 2006CPU benchmark suite, and one from LAPACK---%running on an Intel i7-based machine.  For
%each program, we calculate the permutation entropy of %the processor
%load (instructions per cycle), then compare that to %the prediction accuracy attainable for
%that trace using a series of deterministic models.

The rest of the paper is organized as follows.
Section~\ref{sec:related} discusses previous results on generating
partitions, local modeling, and error distribution analysis, and
situates this work in that context. Section~\ref{sec:methods} covers
the experimental setup and methods used to collect each time
series. Section~\ref{sec:model} describes the prediction models used
in this study.  In Section~\ref{sec:meaComplex}, we review permutation
entropy, the technique that we use to measure complexity.  In
Section~\ref{sec:results}, we estimate the complexity of each time
series and compare that complexity to the accuracy of predictions
produced by the methods of Section~\ref{sec:model}, operating on that
time series.  In Section~\ref{sec:conc}, we discuss these results and
their implications, and consider future areas of research.