### 2014-05-30 pecomplexity.tex
fix an dash

\documentclass[%
pre,
reprint,
superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose,
%preprint,
showpacs,
%preprintnumbers,
nofootinbib,
nobibnotes,
%bibnotes,
 amsmath,amssymb,
 aps,
%pra,
%prb,
%rmp,
%prstab,
%prstper,
%floatfix,
]{revtex4-1}
%\documentclass{article}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
%\geometry{letterpaper}                   % ... or a4paper or a5paper or ...
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parIEEEabrvskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[sort&compress]{natbib}
%\usepackage{caption}
\usepackage{subcaption}
\usepackage{microtype}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\usepackage{amssymb}
%\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage{amsthm}
\usepackage{xspace}
\usepackage{pifont} % check and x marks
\usepackage{color}
\usepackage{listings}
\lstset{language=C}

\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

\newcommand{\gcc}{{\tt 403.gcc}\xspace}
\newcommand{\svd}{{\tt dgesdd}\xspace}
\newcommand{\svdone}{{\tt dgesdd}$_1$\xspace}
\newcommand{\svdtwo}{{\tt dgesdd$_2$}\xspace}
\newcommand{\svdthree}{{\tt dgesdd$_3$}\xspace}
\newcommand{\svdfour}{{\tt dgesdd$_4$}\xspace}
\newcommand{\svdfive}{{\tt dgesdd$_5$}\xspace}
\newcommand{\svdsix}{{\tt dgesdd$_6$}\xspace}
\newcommand{\naive}{na\"ive}
\newcommand{\row}{{\tt row\_major}\xspace}
\newcommand{\col}{{\tt col\_major}\xspace}
\newcommand{\alert}[1]{{\color{red}#1}}
%\newcommand{\color{red}}{{\color[rgb]{1,0,0}}}

%\parskip=3pt

%\usepackage{authblk}
%\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\newtheorem*{mydef}{Definition}

%Version 1
%\author{
%  \IEEEauthorblockN{Joshua Garland}
%  \IEEEauthorblockA{Dept. of Computer Science\\
%    University of Colorado at Boulder\\
%    Colorado, USA\\
%    Email: joshua.garland@colorado.edu}
%  \and
%  \IEEEauthorblockN{Ryan G.~James}
%  \IEEEauthorblockA{Complexity Sciences Center \& Dept. of Physics\\
%    University of California at Davis\\
%    California, USA\\
%    Email: rgjames@ucdavis.edu}
%    \and
%      \IEEEauthorblockN{Elizabeth Bradley}
%         \IEEEauthorblockA{Santa Fe Institute \\
%    New Mexico, USA\\
%    }
%  \IEEEauthorblockA{Dept. of Computer Science\\
%    University of Colorado at Boulder\\
%    Colorado, USA\\
%    Email: lizb@colorado.edu}
%  }

 %\author{
 %  \IEEEauthorblockN{
 %    Joshua Garland\IEEEauthorrefmark{1},
 %    Ryan G.~James\IEEEauthorrefmark{1} and
 %    Elizabeth Bradley\IEEEauthorrefmark{1}\IEEEauthorrefmark{2}}
 %  \IEEEauthorblockA{
 %    \IEEEauthorrefmark{1}Dept. of Computer Science
 %    University of Colorado, Boulder, Colorado 80309-0430 USA\\
 %    Email: joshua.garland@colorado.edu, ryan.james@colorado.edu}
 %  \IEEEauthorblockA{
 %    \IEEEauthorrefmark{2}Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, New Mexico 87501  USA\\
 %    Email: lizb@colorado.edu}
 %  % \IEEEauthorblockA{
 %  %   \IEEEauthorrefmark{3}Complexity Sciences Center \& Physics Dept., University of California, Davis, %California 95616 USA\\
%   %   Email: rgjames@ucdavis.edu}
% }

\begin{document}

\preprint{APS/123-QED}

%\title{Manuscript Title}% Force line breaks with \\
%\thanks{A footnote to the article title}%

\title{
%Determinism, Complexity, and Predictability in Computer Performance\\
%Quantifying structured complexity and its role in predictability: With applications to predicting computer performance.
%\\
Quantifying Time-Series Predictability through Structural Complexity
%\\
%Quantifying Predictability using Structural Complexity
%\\Decoding Predictability in Complicated Real-Valued Time Series: The balance between complexity and temporal structure
%\\Decoding Complicated Time Series: The balance between complexity and temporal structure and the implications for predictability
}

%\author[1]{Joshua Garland %\thanks{joshua.garland@colorado.edu}}
%\author[1]{Ryan James \thanks{ryan.james@colorado.edu}}
%\author[1,2]{Elizabeth Bradley \thanks{lizb@colorado.edu}}
%\affil[1]{Department of Computer Science\\
%  University of Colorado at Boulder\\
%  Colorado, USA
%}
%\affil[2]{Santa Fe Institute\\
%  New Mexico, USA
%}

\author{Joshua Garland}
\email{joshua.garland@colorado.edu}
\affiliation{Department of Computer Science, University of Colorado at Boulder. Colorado, USA}

\author{Ryan James}
\email{ryan.james@colorado.edu}
\affiliation{Department of Computer Science, University of Colorado at Boulder. Colorado, USA}

\author{Elizabeth Bradley}
\email{lizb@colorado.edu}
\affiliation{Department of Computer Science, University of Colorado at Boulder. Colorado, USA}
\affiliation{Santa Fe Institute, New Mexico, USA}

\date{\today}% It is always \today, today,
             %  but any date may be explicitly specified

\begin{abstract}

This paper provides insight into when, why, and how forecast
strategies fail when they are applied to complicated time series. We
conjecture that the inherent complexity of a time series --- which
results from the dimension, nonlinearity, and non-stationarity of the
generating process, as well as from measurement issues like noise,
aggregation, and finite data length --- is both empirically
quantifiable and directly correlated with predictability.  In
particular, we argue that \emph{redundancy} is an effective way to
measure complexity and predictive structure in a time series and that
\emph{weighted permutation entropy} is an effective way to estimate
that complexity.  To validate these conjectures, we study 120
different time-series data sets.  For each time series, we construct
predictions using a wide variety of forecast models, then compare the
accuracy of the predictions with the permutation entropy of that time
series.  We use the results to develop a heuristic that can help
practitioners recognize when a particular prediction method is not
well matched to the task at hand: that is, when the time series has
more predictive structure than that method can capture and exploit.

\vspace{0.1in}
\noindent
{\bf Keywords}: entropy, permutation entropy, weighted permutation
entropy, time-series analysis, predictability

\preprint{Santa Fe Institute Working Paper XX-XX-XXX}
\preprint{arxiv.org:XXXX.XXXX [nlin.CD]}

\end{abstract}
\pacs{
05.45.Tp    % Time series analysis (nonlinear dynamics)
89.70.-a    % Information theory
%89.70.Hj    %   communication complexity
%89.70.Cf    %   entropy
%05.45.-a    % Nonlinear dynamics
89.75.Kd    %   Patterns (complex systems)
%95.75.Wx    % Time series analysis, time variability
95.75.Pq    % Mathematical procedures and computer techniques
%07.05.-t    % Computers in experimental physics
}

\maketitle

\input{intro.tex}
\input{related.tex}
\input{methods.tex}

\input{model.tex}

\input{wpe.tex}

\input{results.tex}

\input{concl.tex}

\section*{Acknowledgment}
This work was partially supported by NSF grant \#CMMI-1245947 and ARO
grant \#W911NF-12-1-0288.
We would also like to acknowledge the Santa Fe Institute whose interdisciplinary halls created a place for this work to be born.

%\bibliographystyle{aipauth4-1}
\bibliographystyle{apsrev4-1}
\bibliography{bibliofile}

\end{document}